---
title: "#SundayStats: A deep dive into the stats package in R - The Normal Distribution"
author: 'Jose Francisco Endrinal'
output: github_document
---

*DISCLAIMER: I am not a working data analyst, nor am I a data scientist. As such, the code, techniques, and methods used in this blog post do not qualify as industry-level code, techniques, and methods. This blog and its corresponding github repository are meant to document my progress as I come to learn the techniques and skills a data analyst or data scientist will be needing in their line of work. This style of learning is popularly called "Learning in Public"*

## Introduction

*\#SundayStats is a weekly post on this blog about the default `stats` package in R. Each week (usually when I feel like it), I will be working with one function/group of functions in this package and will talk about how to use it in a practical setting.*  

If you've taken a statistics class, you probably know about some fancy concept called the normal distribution. If you're like most people and myself, you would have forgetten its importance the second you graduate from university. While most fields are doing statistical analysis as a way to test their claims about the world, the backbone of most of that analysis is often misunderstood or taken for granted. Yes, the normal distribution underpins most if not all scientific and social science quantitatively-backed study, including useful tools they use like linear regression, hypothesis testing, and the like. And yet most of us do statistical work blissfuly unaware of its properties and the role they play in our analysis.  

This blog post will try to help in the reader's understanding of the normal distribution by demonstrating how it's used in the R `stats` package. I won't pretend this will be an exhaustive blog post, but it will broadly cover the concept enough to help the reader grasp its characteristics and understand its use in statistical analysis.  

But first, if there is such a thing a normal distribution. What is a distribution in the first place?  

As a quick review, you know that when you want to understand how long, how short, how many or how few of a thing in the world is, you collect enough data and begin to realize that quantities/dimensions of things tend to not just be fixed at one value. What you'll find is that the measurements usually center around a number somewhere, and that generally, the measures/counts will spread a certain way.  

A distribution is a model that describes how frequently countable increasing/decreasing numbers would occur given that your randomly sampled data gets very large (massive oversimplification here, trying to talk to a general audience). Put simply, it is a claim about whether the data is centered or spread a certain way and and attempt to quantify how much.  

Now a normal distribution is a special distribution where most of the values are near a "center" and fewer of the values are farther away from the center. Now a lot of other types of distributions can have this characteristic, but most people would be happy with saying that something is normally distributed. You can argue about whether some other distribution better represents the data on hand. But because the normal distribution is more familiar to work with from a statistical standpoint, people tend to use it.   

Let's take a deep dive. 

```{r}
suppressPackageStartupMessages({
  library(tidyverse)})
```

## The function definition of the normal distribution

Here is the formula for the normal distribution:

```{r}
norm_dist <- function(x, mu, sigma) {
  e**-(x)
}
```


## Using rnorm() to generate a normally distributed dataset

```{r}
# Generate dataset
set.seed(2387)
df <- rnorm(n = 150, mean = 0, sd = 1)
```

Let's plot df:  

```{r}
ggplot(data_frame(df = df), aes(df)) + 
  geom_density(fill = "#4529A3", color = "#4529A3", alpha = 0.25)
```

Somewhat normal. Good enough.  

## Using pnorm() to generate a normally distributed function with quantiles

```{r}
# show function/s with all their attributes
df <- pnorm(c(-2.75, -1.25, 0, 1.25, 2.75))
```

## Output

```{r}
# Visualize/print output
```

***
More dataset information: [[Dataset description]](link to website)  

If you liked this blog post, feel free to give me a clap!  

If you'd like to give feedback, you can reach out to me at:  
Email: francis.endrinal@gmail.com  
FB Messenger: m.me/transparencyman  
Twitter, Instagram: @jgendrinal
