---
title: "Back to Iris: testing in R"
author: 'Jose Francisco Endrinal'
output: github_document
---

*DISCLAIMER: I am not a working data analyst, nor am I a data scientist. As such, the code, techniques, and methods used in this blog post do not qualify as industry-level code, techniques, and methods. This blog and its corresponding github repository are meant to document my progress as I come to learn the techniques and skills a data analyst or data scientist will be needing in their line of work. This style of learning is popularly called "Learning in Public"*  

## Introduction

Testing your code seems to be an important part of the data analysis process. For one, every step of processing data works under certain assumptions. It's not always the case that those assumptions are true. And when your assumptions are not in check, the steps in your analysis break down.  

Another reason to test your code is that models also have the same assumption. In linear regression alone, there are five major assumptions that you have to keep in order to the results of linear regression to be unbiased. Testing your code ensures that the model fulfills all those assumptions.  

## Setup

```{r}
# Load packages
suppressPackageStartupMessages({
  library(tidyverse)
  library(testthat) # for unit tests
  library(assertthat)}) # for function tests
# Load dataset
read_csv("data/iris.csv") -> iris.dt
```

## First look at the dataset

```{r}
# View dataset
```


## Exploratory Data Analysis

```{r}
# Distribution of the class/target over the variable

```





***
More dataset information: [[Dataset description]](link to website)  

Feedback:  
Email: francis.endrinal@gmail.com  
FB Messenger: m.me/transparencyman  
Twitter, Instagram: @jgendrinal
